---
title: "Enhancing {manifestoR} data objects: A short tutorial using the Swiss manifesto corpus."
author: "Hauke Licht"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Short tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE
  , eval = TRUE
  , warning = FALSE
  , message = FALSE
  , cache = FALSE
  , fig.align = 'center'
  , fig.height = 4
  , fig.width = 8
)
```

```{r packages}
library(manifestoR)
library(manifestoEnhanceR)
library(dplyr)
library(tidyr)
library(purrr)
library(xml2)
library(rvest)
```

# Step 1: load manifesto data 

```{r set_api_key}
# set Manifesto Project API key
mp_setapikey(file.path(Sys.getenv("SECRETS_PATH"), "manifesto_apikey.txt")) 
```

```{r get_cmp_data}
# get CMP party-year data
cmp <- mp_maindataset(south_america = FALSE)

# subset to Swiss configurations
ch_configs <- cmp %>% 
  filter(countryname == "Switzerland") %>% 
  select(1:10)

# get Swiss manifestos
ch <- mp_corpus(countryname == "Switzerland", cache = TRUE)
```

## Conversion to tidy data

```{r as_tibble}
# this is a 'ManifestoCorpus' object
class(ch)
# it's a list 'ManifestoDocument' objects
str(head(ch, 2), 1)
lapply(head(ch, 2), class)

# we can tidy this up (i.e., convert it to a long tibble)
man_dfs <- as_tibble(ch)
```

Column 'data' is a list column, nesting manifesto tibbles. 

```{r add_meta_data}
# left join manifesto- and party-level data 
man_sents <- man_dfs %>% 
  left_join(
    ch_configs %>% 
      transmute(manifesto_id = paste0(party, "_", date), party, partyname, partyabbrev) %>% 
      unique()
  )
```

# Step 2: Define some test cases

```{r def_test_cases}
test1_df <- unnest(filter(man_sents, manifesto_id == "43120_201110"), data)
test2_df <- unnest(filter(man_sents, manifesto_id == "43110_198710"), data)
test3_df <- unnest(filter(man_sents, manifesto_id == "43110_199910"), data)
test4_df <- unnest(filter(man_sents, manifesto_id == "43120_201510"), data)
```

# Step 3: Enhance manifesto data frames

By enhancing manifesto data frames, we add bloc, sentence and quasi-sentence (qs) counters, as well as a role indicators ('qs', 'header', 'title', or 'meta').

```{r enhance_test_1}
test1_res <- enhance_manifesto_df(test1_df)
select(test1_res, 1:4, 17:25)
class(test1_res)
attr(test1_res, "annotated")
attr(test1_res, "extra_cols")
"title" %in% test1_res$role
```

```{r enhance_test_2}
test2_res <- enhance_manifesto_df(test2_df)
select(test2_res, 1:4, 17:25)
class(test2_res)
attr(test2_res, "annotated")
```

```{r enhance_test_3}
test3_res <- enhance_manifesto_df(test3_df)
class(test3_res)
attr(test3_res, "annotated")
"title" %in% test3_res$role
```

```{r enhance_test_4}
test4_res <- enhance_manifesto_df(test4_df)
class(test4_res)
attr(test4_res, "annotated")
"title" %in% test4_res$role
```

# Step 4: Convert manifesto data frames to XML documents

```{r xmlify_test_1}
# test case 1: as string
test1_xml <- manifesto_df_to_xml(test1_res, parse = FALSE)
class(test1_xml)
length(test1_xml)
test1_xml <- manifesto_df_to_xml(filter(test1_res, bloc_nr < 3), parse = FALSE)
# cat(test1_xml)

# as XML document
test1_xml <- manifesto_df_to_xml(test1_res)
test1_xml <- manifesto_df_to_xml(man.df = filter(test1_res, bloc_nr %in% 9:9))
class(test1_xml)
test1_xml
```

```{r xmlify_test_2}
# test case 2
test2_xml <- manifesto_df_to_xml(test2_res, parse = FALSE)
cat(test2_xml)
test2_xml <- manifesto_df_to_xml(test2_res)
class(test2_xml)
test2_xml
```

```{r xmlify_test_3}
# test case 3
test3_xml <- manifesto_df_to_xml(man.df = test3_res, parse = FALSE)
class(test3_xml); length(test3_xml)
test3_xml <- manifesto_df_to_xml(test3_res)
class(test3_xml)
test3_xml
html_node(test3_xml, "title")
```

```{r xmlify_test_4}
# test case 4
test4_xml <- manifesto_df_to_xml(man.df = test4_res, parse = FALSE)
class(test4_xml); length(test4_xml)
test4_xml <- manifesto_df_to_xml(test4_res)
class(test4_xml)
test4_xml
html_node(test4_xml, "title")
```

# Step 5: Apply to all manifestos in one pipeline

```{r parse_all_manifestos}
man_xmls <- man_sents %>% 
  # drop unneccessary columns
  select(-has_eu_code, -may_contradict_core_dataset, -md5sum_text, -md5sum_original, -annotations, -id) %>% 
  # unnest data
  unnest(data) %>% 
  # split by manifesto ID (-> list od DFs)
  split(.$manifesto_id) %>% 
  # apply enhance
  map(enhance_manifesto_df) %>% 
  # apply converter
  map(safely(manifesto_df_to_xml)) %>% 
  # gather all in list-column data frame 
  enframe() %>% 
  transmute(
    manifesto_id = name
    , xml = map(value, "result")
    , errors = map(map(value, "error"), "message")
    , n_errors = lengths(errors)
  )

# the returned tibble contains xml_documents in the 'xml' list-column
head(man_xmls, 3)

# there were no errors raised
man_xmls %>% 
  filter(n_errors > 0) %>% 
  unnest(errors)
```

# Step 6 (optional): write manifestos to disk

```{r , eval=FALSE}
# write to disk
map2(
  man_xmls$xml
  , man_xmls$manifesto_id
  , function(xml, id, path = file.path("your", "local", "path")) {
  write_xml(xml, file.path(path, paste0(id, ".xml")))  
})
```

